---
title: "Comparison of varianace in Correspondance analysis and Beta diversity rewiring contribution"
author:
  - name: "Rémi Legrand"
  - name: "Second Author's name"
abstract: >
  The goal is to look at the variance of the positions in the Correspondance Analysis that we think is a proxy of the ability of one specie to do rewiring. And the corruent way of quantifying the individual contribution to rewiring which was defined in Toju et al (2024), consisting in using the metrics of beta diversity defined by Frund or the other (to check!!!) and assume that by removing the specie and computing again the Beta diversities, we can obtain the contribution of the specie.
  
  This would significantly fastened the computation if it works.
date: "`r format(Sys.time(), '%d %B %Y')`"
url: https://GitHubID.github.io/Repository/
github-repo: GitHubID/Repository
# Language
lang: en-US
otherlangs: [fr-FR,it]
preamble: >
  \hyphenation{bio-di-ver-si-ty sap-lings}
# Print table of contents in PDFs?
pdftoc: false
# If true, choose its depth
toc-depth: 3
# URL color
urlcolor: blue
# Do not modify
always_allow_html: yes
csquotes: true
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    css: style.css
    code_folding: show
  rmdformats::downcute:
    use_bookdown: yes
    lightbox: yes
  bookdown::gitbook:
    config:
      download: "pdf"
      sharing:
        github: yes
  bookdown::pdf_book:
    template: latex/template.tex
    citation_package: natbib
    latex_engine: xelatex
    keep_tex: yes
  bookdown::word_document2: default
---

# Compare Variance to beta diversity contribution

```{r message=FALSE, warning=FALSE}
library(NetworkEnvGradGeneration)
library(ade4)
library(stringr)
library(igraph)
library(bipartite)
library(parallel)
library(ggplot2)
```

## Extract the giant component

### Extract the components that are not always part of the giant component

```{r}
connected_compo <- function(simu_netw, nb_frame){
  not_always_con <- c() # to store the non connected species
  sampled_netw <- Reduce('+', simu_netw$obs_network) # overlap of all the networks
  sampled_netw_con = sampled_netw[which(rowSums(sampled_netw) != 0), which(colSums(sampled_netw) != 0)] # all components that have at least one connection
  component <- igraph::components(graph_from_incidence_matrix(sampled_netw_con))
  if (length(component$csize) != 1) {
      not_always_con <- c(not_always_con, names(component$membership[which(component$membership != which.max(component$csize))]))
  }
  return(unique(not_always_con))
}
```

```{r}
filter_giant_component <- function(simu_netw, not_giant_compo){
  # On observed data
  simu_netw$obs_giant_compo <- lapply(simu_netw$obs_network, function(arr){
    arr[!dimnames(arr)[[1]] %in% unique(not_giant_compo),
        !dimnames(arr)[[2]] %in% unique(not_giant_compo)]
  }) # keep only the data that is not inside of the list of the non connected components
  
  # On the trait matching matrix
  simu_netw$trait$trait_mat_giant_compo <- simu_netw$trait$matching_matrix[!dimnames(simu_netw$obs_network[[1]])[[1]] %in% unique(not_giant_compo),
                                                                           !dimnames(simu_netw$obs_network[[1]])[[2]] %in% unique(not_giant_compo)]
  
  # On the resource traits
  simu_netw$trait$resource_giant_compo <- simu_netw$trait$trait_resource[!dimnames(simu_netw$trait$trait_resource)[[1]] %in% unique(not_giant_compo),]
  
  #On the consumer traits
  simu_netw$trait$consumer_giant_compo <- simu_netw$trait$trait_consumer[!dimnames(simu_netw$trait$trait_consumer)[[1]] %in% unique(not_giant_compo),,]
  
  return(simu_netw)
}
```

## Simulate the data

```{r}
simu_network <- env_grad_netw(nb_resource = 50, nb_consumer = 50, nb_location = 10,
              ninter = 40000,
              mean_tol_env = 0.7, mean_tol = 0.1,
              ratio_grad = 0.5, delta = 0.2)

############### Filter to only keep the giant component ###############
not_giant_compo <- connected_compo(simu_network, nb_frame)
simu_network <- filter_giant_component(simu_network, not_giant_compo)
tab = Reduce('+', simu_network$obs_giant_compo) # afin de générer le réseau moyen

############### AFC sur le réseau moyen ###############
afc.tab=dudi.coa(tab, nf = 6, scannf = FALSE)
scatter(afc.tab)

############### AFC su chacun des réseaux en utilisant les axes précédents ###############
fou1 = foucart(lapply(simu_network$obs_network, as.data.frame), scannf = F, nf = 2)
```

```{r}
fou1$eig[1]
```

## Variance position in the Correspondance analysis

Here, we will need to fuse the two axis because we only have one values for the beta diversity. Hence we can fuse two of them by doing the sum of the variance on the axis time the eigen-value of the corresponding axis.

```{r}
library(stringr)

############### Initiate the arrays ###############
res_var <- array(NA, c(nrow(tab), 3))
dimnames(res_var)[[1]] <- rownames(tab)
dimnames(res_var)[[2]] <- c("axis1", "axis2", "tot_res")

con_var <- array(NA, c(nrow(tab), 3))
dimnames(con_var)[[1]] <- colnames(tab)
dimnames(con_var)[[2]] <- c("axis1", "axis2", "tot_con")

############### Compute variance ###############

######## AXIS 1 ########
# for resources
res_var_ax1 <- c()
for (res in 1:nrow(tab)) {
  res1 <- fou1$Tli[str_detect(string = dimnames(fou1$Tli)[[1]],
                              pattern = paste0("^res", res, ".Ana")), ]
  res_var_ax1 <- c(res_var_ax1, sd(res1[, 1]))
}
hist(res_var_ax1, breaks = 20)
res_var[, 1] <- res_var_ax1

# for consumers 
con_var_ax1 <- c()
for (con in 1:ncol(tab)) {
  con1 <- fou1$Tco[str_detect(string = dimnames(fou1$Tco)[[1]],
                              pattern = paste0("^con", con, ".Ana")), ]
  con_var_ax1 <- c(con_var_ax1, sd(con1[, 1]))
}
hist(con_var_ax1, breaks = 20)

######## AXIS 2 ########
# for resources
res_var_ax2 <- c()
for (res in 1:nrow(tab)) {
  res1 <- fou1$Tli[str_detect(string = dimnames(fou1$Tli)[[1]],
                              pattern = paste0("^res", res, ".Ana")), ]
  res_var_ax2 <- c(res_var_ax2, sd(res1[, 2]))
}
hist(res_var_ax2, breaks = 20)

# for consumers 
con_var_ax2 <- c()
for (con in 1:ncol(tab)) {
  con1 <- fou1$Tco[str_detect(string = dimnames(fou1$Tco)[[1]],
                              pattern = paste0("^con", con, ".Ana")), ]
  con_var_ax2 <- c(con_var_ax2, sd(con1[, 2]))
}
hist(con_var_ax2, breaks = 20)

############### Tidy variance in dataframe ###############
res_var[, 1] <- res_var_ax1
con_var[, 1] <- con_var_ax1
res_var[, 2] <- res_var_ax2
con_var[, 2] <- con_var_ax2

############### Merge information from both axis ###############

res_var[ , 3] <- res_var_ax1 * fou1$eig[1] + res_var_ax2 * fou1$eig[2]
con_var[ , 3] <- con_var_ax1 * fou1$eig[1] + con_var_ax2 * fou1$eig[2]

hist(res_var[, 3], breaks = 20)
hist(con_var[, 3], breaks = 20)
```

## Beta diversity contribution

Here I first look at the beta-diversity between two frames for each possible combination of frames.

```{r}
array_3d <- array(unlist(simu_network$obs_network),
                  dim = c(nrow(simu_network$obs_network[[1]]),
                          ncol(simu_network$obs_network[[1]]),
                          length(simu_network$obs_network)))

dimnames(array_3d)[[1]] <- rownames(simu_network$obs_network[[1]])
dimnames(array_3d)[[2]] <- colnames(simu_network$obs_network[[1]])
dimnames(array_3d)[[3]] <- paste0("frame",1:length(simu_network$obs_network))
# array_3d
```

```{r}
(nw_beta_div <- betalinkr_multi(array_3d, partitioning="poisot", binary=FALSE))

mean(nw_beta_div$OS)
mean(nw_beta_div[-1,"OS"])
mean(betalinkr_multi(array_3d, partitioning="poisot", binary=FALSE)[-17,"OS"])
```

### For resources

here the goal is to remove one species at the time and then compute mean the multi beta diversity for each possible combination of frames $\Delta\beta_{OS,i} = mean(\Delta\beta_{OS,i,f_a,f_b})$.

```{r}
start_time <- Sys.time()

beta_indiv_res <- array(data = NA, 
                    dim = c(nrow(simu_network$obs_network[[1]]), 1))

rownames(beta_indiv_res) = rownames(simu_network$obs_network[[1]])
colnames(beta_indiv_res) = "delta_os"
# beta_indiv

my_function <- function(x) {
  mean(betalinkr_multi(aperm(array_3d, c(2,3,1)), partitioning="poisot", binary=FALSE)[-x, "OS"])
}

# Define a list of input values
input_values <- 1:nrow(simu_network$obs_network[[1]])

# Run the computations in parallel using mclapply
# number of cores :'mc.cores'
result <- mclapply(input_values, my_function, mc.cores = 4)

# Print the results
beta_indiv_res[,1] <- unlist(result)

end_time <- Sys.time()
end_time - start_time # running time
hist(mean(nw_beta_div$OS) - beta_indiv_res[, 1], breaks = 10) # gives the distance to the mean contribution of the species
hist(beta_indiv_res[, 1], breaks = 10) # same but with only the contribution to rewiring for each species
```

```{r}
# mean(betalinkr_multi(aperm(array_3d, c(1,3,2)), partitioning="poisot", binary=FALSE)[-x, "OS"])
```


```{r}
# aperm(array_3d, c(2,3,1))
```

```{r}
mean(beta_indiv_res[, 1])
mean(nw_beta_div$OS)
```


### For consumers

```{r}
start_time <- Sys.time()

beta_indiv_con <- array(data = NA, 
                    dim = c(ncol(simu_network$obs_network[[1]]), 1))

rownames(beta_indiv_con) = colnames(simu_network$obs_network[[1]])
colnames(beta_indiv_con) = "delta_os"
# beta_indiv

# Define a function that simulates some computational work
my_function <- function(x) {
  mean(betalinkr_multi(aperm(array_3d, c(1,3,2)), partitioning="poisot", binary=FALSE)[-x, "OS"])
}

# Define a list of input values
input_values <- 1:nrow(simu_network$obs_network[[1]])

# Run the computations in parallel using mclapply
# Here, we specify the number of cores to be used with the 'mc.cores' argument
# You can adjust this number based on the number of cores available on your machine
result <- mclapply(input_values, my_function, mc.cores = 4)

# Print the results
beta_indiv_con[,1] <- unlist(result)

end_time <- Sys.time()
end_time - start_time

br <- 8
hist(mean(nw_beta_div$OS) - beta_indiv_con[, 1], main = "distance to beta_os",breaks = br)
hist(mean(beta_indiv_con[, 1]) - beta_indiv_con[, 1], main = "distance to mean delta_beta_os", breaks = br)
hist(- beta_indiv_con[, 1], main = "contribution to beta_os (delta_beta_os)",breaks = br)
```

```{r}
sum(mean(nw_beta_div$OS) - beta_indiv_con[, 1])
mean(nw_beta_div$OS)
```


## Comparison between the variance and the beta diversity

```{r}
ggplot() +
  geom_point(data = as.data.frame(cbind(con_var, beta_indiv_con)), aes(x = tot_con, y = delta_os)) +
  xlab("variance in the correspondance analysis") +  # Add X axis label
  ylab("delta OS indiv") +  # Add Y axis label
  ggtitle("Comparison of the varinace in trhe correspondance analysis \n against the individual contribution to beta os for consumers") +  # Add plot title
  theme_bw()

ggplot() +
  geom_point(data = as.data.frame(cbind(res_var, beta_indiv_res)), aes(x = tot_res, y = delta_os)) +
  xlab("variance in the correspondance analysis") +  # Add X axis label
  ylab("delta OS indiv") +  # Add Y axis label
  ggtitle("Comparison of the varinace in trhe correspondance analysis \n against the individual contribution to beta os for resources") +  # Add plot title
  theme_bw()
```



