---
title: "reconstruct_trait_simu"
author: "Rémi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(NetworkEnvGradGeneration)
library(ade4)
library(igraph)
```


```{r}
##################################################"
## reconst.coa reconstitue le tableau original Nij
## ici on reconstitue les ecarts entre pij observé et théorique pi.p.j
## (pij - pi.p.j) pour reconst_afc ou normés par le chi2 ((pij - pi.p.j)/pi.p.j)) pour reconst_afc2
##################################################"

library(ade4)

reconst_afc <- function (dudi, nf = 1, ...) 
{
    ## reconst (pij - pi.p.j)/pi.p.j
    if (!inherits(dudi, "dudi")) 
        stop("Object of class 'dudi' expected")
    if (nf > dudi$nf) 
        stop(paste(nf, "factors need >", dudi$nf, "factors available\n"))
    if (!inherits(dudi, "coa")) 
        stop("Object of class 'dudi' expected")
    pl <- dudi$lw
    pc <- dudi$cw
    n <- dudi$N
    #res0 <- outer(pl, pc) * n
    res0 <- matrix(1, length(pl), length(pc)) # c'est bien pc?
    res <- data.frame(res0-1) # pourquoi pas direct matr(0, pl,pc)?
    names(res) <- names(dudi$tab) # col.names?
    row.names(res) <- row.names(dudi$tab)
    #res0 <- outer(pl, pc) * n
    if (nf == 0) 
        return(res)
    for (i in 1:nf) {
        xli <- dudi$li[, i]
        yc1 <- dudi$c1[, i] # pourquoi li et C1 et pas l1/c1, ou li/co? 
        res <- res + outer(xli, yc1) * res0
    }
    return(res)
} 


reconst_afc2 <- function (dudi, nf = 1, ...) 
{
    ## reconst (pij - pi.p.j)
    if (!inherits(dudi, "dudi")) 
        stop("Object of class 'dudi' expected")
    if (nf > dudi$nf) 
        stop(paste(nf, "factors need >", dudi$nf, "factors available\n"))
    if (!inherits(dudi, "coa")) 
        stop("Object of class 'dudi' expected")
    pl <- dudi$lw
    pc <- dudi$cw
    n <- dudi$N
    res0 <- outer(pl, pc) * n
    res <- data.frame(res0 - 1)
    names(res) <- names(dudi$tab)
    row.names(res) <- row.names(dudi$tab)
    if (nf == 0) 
        return(res)
    for (i in 1:nf) {
        xli <- dudi$li[, i]
        yc1 <- dudi$c1[, i]
        res <- res + outer(xli, yc1) * res0
    }
    return(res)
}
```

# Generate Network

```{r}
nb_frame = 16
spe_resource = 70
spe_consumer = 190
simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                   nb_location = nb_frame,
                                   know_env_grad_pos = F, ninter = 2400, 
                                   mean_tol = 0.2, sd_tol = 0.2,
                                   mean_tol_env = 100, sd_tol_env = 0.3,
                                   delta = 1, corr_env_trait = 0)

```


## Extract the giant component

### Extract the components that are not always part of the giant component

```{r}
connected_compo <- function(simu_netw, nb_frame){
  not_always_con <- c()
  sampled_netw <- Reduce('+', simulated_network$obs_network)
  sampled_netw_con = sampled_netw[which(rowSums(sampled_netw) != 0), which(colSums(sampled_netw) != 0)]
  component <- igraph::components(graph_from_incidence_matrix(sampled_netw_con))
  if (length(component$csize) != 1) {
      not_always_con <- c(not_always_con, names(component$membership[which(component$membership != which.max(component$csize))]))
  }
  
  # print(paste(length(unique(not_always_con))/(spe_consumer + spe_resource) * 100,"% of the species will be removed because there is some frames where they are not part of the giant component"))
  return(unique(not_always_con))
}
# sampled_netw
not_giant_compo <- connected_compo(simulated_network, nb_frame)
```

```{r}
filter_giant_component <- function(simu_netw, not_giant_compo){
  # On observed data
  simu_netw$obs_giant_compo <- lapply(simu_netw$obs_network, function(arr){
    arr[!dimnames(arr)[[1]] %in% unique(not_giant_compo),
        !dimnames(arr)[[2]] %in% unique(not_giant_compo)]
  })
  
  # On the trait matching matrix
  simu_netw$trait$trait_mat_giant_compo <- simu_netw$trait$matching_matrix[!dimnames(simu_netw$obs_network[[1]])[[1]] %in% unique(not_giant_compo),
                                                                                !dimnames(simu_netw$obs_network[[1]])[[2]] %in% unique(not_giant_compo)]
  
  # On the resource traits
  simu_netw$trait$resource_giant_compo <- simu_netw$trait$trait_resource[!dimnames(simu_netw$trait$trait_resource)[[1]] %in% unique(not_giant_compo),]
  
  #On the consumer traits
  simu_netw$trait$consumer_giant_compo <- simu_netw$trait$trait_consumer[!dimnames(simu_netw$trait$trait_consumer)[[1]] %in% unique(not_giant_compo),,]
  
  return(simu_netw)
}

simulated_network <- filter_giant_component(simulated_network, not_giant_compo)

# not_giant_compo
# simulated_network$trait$trait_mat_giant_compo

```


```{r}
ade4::table.value(simulated_network$obs_giant_compo[[1]])
```

```{r}
sampled_netw <- Reduce('+', simulated_network$obs_giant_compo)
# to keep only the connected species
sampled_netw_con <- sampled_netw[which(rowSums(sampled_netw) != 0), which(colSums(sampled_netw) != 0)]
component <- igraph::components(graph_from_incidence_matrix(sampled_netw_con))
if (length(component$csize) != 1) {
  print(paste("The network",i, "is not strongly connected, there is", component$no, "connected components"))
  print(names(component$membership[which(component$membership != which.max(component$csize))]))
  print("are not part of of the giant component.")
  not_always_con <- c(not_always_con, names(component$membership[which(component$membership != which.max(component$csize))]))
}
table.value(sampled_netw)

plot(graph_from_incidence_matrix(sampled_netw), vertex.size=5)
```

```{r}
degree(graph_from_incidence_matrix(sampled_netw))
```


```{r}
tab = Reduce('+', simulated_network$obs_giant_compo) # afin de générer le réseau moyen

afc.tab=dudi.coa(tab, nf = 18, scannf = FALSE)

par(mfrow=c(2,2))
ade4::table.value(simulated_network$trait$trait_mat_giant_compo, x = rank(afc.tab$co[,1]), y = rank(afc.tab$li[,1]), csize = 0.8) # trait matching matrix et tri selon le premier axe
ade4::table.value(tab, x = rank(afc.tab$co[,1]), y = rank(afc.tab$li[,1])) # réseau moyen observé
toto = reconst_afc(afc.tab, nf = 6) # pour reconstruire le réseau moyen à partir des positions sur l'AFC mais on voudrait plustôt obtenir le trait matching
ade4::table.value(toto, x = rank(afc.tab$co[,1]), y = rank(afc.tab$li[,1]), csize = 0.8) # plot le réseau reconstruit
ade4::table.value(afc.tab$tab, x = rank(afc.tab$co[,1]), y = rank(afc.tab$li[,1])) # transformed average table
```

```{r}
afc.tab=dudi.coa(tab, nf = 18, scannf = F)
```


```{r}
# toto = reconst_afc(afc.tab, nf = 4)
# plot(RV.randtest(toto, data.frame(simulated_network$trait$trait_mat_giant_compo), nrepet = 999))
# 
# rv <- RV.randtest(toto, data.frame(simulated_network$trait$trait_mat_giant_compo), nrepet = 999)
# 
# rv$obs
```

```{r}
library(ggplot2)
library(tidyr)
```

```{r}
toto = reconst_afc(afc.tab, nf = 12) # pour reconstruire le réseau moyen à partir des positions sur l'AFC mais on voudrait plustôt obtenir le trait matching
toto
toto$res <- rownames(toto)
reconstructed = pivot_longer(toto, cols = starts_with("con"), names_to = "con")
reconstructed$res <- as.factor(reconstructed$res)
reconstructed$con <- as.factor(reconstructed$con)
str(reconstructed)

# hist(reconstructed$value)

# reconstructed$value[which(reconstructed$value < 0)] <- 0

ggplot(reconstructed, aes(x = res, y = con, fill = value)) + geom_tile() + scale_fill_gradient(low = "white", high = "blue")
```

```{r}
# ade4::table.value(simulated_network$trait$trait_mat_giant_compo[rank(afc.tab$li[,1]), rank(afc.tab$co[,1])])
# rank(afc.tab$li[,1])

trait_mat <- as.data.frame(simulated_network$trait$trait_mat_giant_compo)
trait_mat$res <- rownames(simulated_network$trait$trait_mat_giant_compo)
trait_mat <- pivot_longer(trait_mat, cols = starts_with("con"), names_to = "con")
trait_mat$res <- as.factor(trait_mat$res)
trait_mat$con <- as.factor(trait_mat$con)

ggplot(trait_mat, aes(x = res, y = con, fill = value)) + geom_tile() + scale_fill_gradient(low = "white", high = "blue")
```


```{r}
# afc.tab$li

cor(simulated_network$trait$resource_giant_compo, afc.tab$li) # resource correlation with the rows of the CA depending on the axis
cor(simulated_network$trait$consumer_giant_compo[,1,], afc.tab$co) # consumer correlation with the columns of the CA depending on the axis
```

```{r}
scatter(afc.tab)
```


```{r}
library(adegraphics)
adegraphics::table.value(afc.tab$tab, coordsx = rank(afc.tab$co[,1]), coordsy  = rank(afc.tab$li[,1]), method = "color", ppoint.cex = 0.2) # transformed average table

image(simulated_network$trait$matching_matrix [order(afc.tab$li[,1]), order(afc.tab$co[,1])])

```


```{r}
tr1x1_line <- c()
tr1x2_line <- c()
tr1x3_line <- c()
tr1x4_line <- c()

tr2x1_line <- c()
tr2x2_line <- c()
tr2x3_line <- c()
tr2x4_line <- c()

tr1x1_col <- c()
tr1x2_col <- c()
tr1x3_col <- c()
tr1x4_col <- c()

tr2x1_col <- c()
tr2x2_col <- c()
tr2x3_col <- c()
tr2x4_col <- c()

for (i in 1:5) {
  nb_frame = 5
  spe_resource = 50
  spe_consumer = 100
  simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                     nb_location = nb_frame,
                                     know_env_grad_pos = F, ninter = 100, 
                                     mean_tol = 0.1, sd_tol = 0.1,
                                     mean_tol_env = 0.1, sd_tol_env = 0.1,
                                     delta = 1, corr_env_trait = 0,
                                     ratio_grad = 0.6)
  
  not_giant_compo <- connected_compo(simulated_network, nb_frame)
  simulated_network <- filter_giant_component(simulated_network, not_giant_compo)
  
  
  tab = Reduce('+', simulated_network$obs_giant_compo) # afin de générer le réseau moyen
  afc.tab=dudi.coa(tab, nf = 18, scannf = FALSE)
  
  (corr_line <- cor(simulated_network$trait$resource_giant_compo, afc.tab$li)) # resource correlation with the rows of the CA depending on the axis
  (corr_col <- cor(simulated_network$trait$consumer_giant_compo[,1,], afc.tab$co)) # consumer correlation with the columns of the CA depending on the axis
  
  tr1x1_line <- c(tr1x1_line, abs(corr_line[1,1]))
  tr1x2_line <- c(tr1x2_line, abs(corr_line[1,2]))
  tr1x3_line <- c(tr1x3_line, abs(corr_line[1,3]))
  tr1x4_line <- c(tr1x4_line, abs(corr_line[1,4]))
  
  tr2x1_line <- c(tr2x1_line, abs(corr_line[2,1]))
  tr2x2_line <- c(tr2x2_line, abs(corr_line[2,2]))
  tr2x3_line <- c(tr2x3_line, abs(corr_line[2,3]))
  tr2x4_line <- c(tr2x4_line, abs(corr_line[2,4]))

  tr1x1_col <- c(tr1x1_col, abs(corr_col[1,1]))
  tr1x2_col <- c(tr1x2_col, abs(corr_col[1,2]))
  tr1x3_col <- c(tr1x3_col, abs(corr_col[1,3]))
  tr1x4_col <- c(tr1x4_col, abs(corr_col[1,4]))
  
  tr2x1_col <- c(tr2x1_col, abs(corr_col[2,1]))
  tr2x2_col <- c(tr2x2_col, abs(corr_col[2,2]))
  tr2x3_col <- c(tr2x3_col, abs(corr_col[2,3]))
  tr2x4_col <- c(tr2x4_col, abs(corr_col[2,4]))
  
  s.corcircle(corr_line)
  s.corcircle(corr_col)
  scatter(afc.tab)
}
# par(mfrow = c(2,1))
boxplot(tr1x1_line, tr1x2_line, tr1x3_line, tr1x4_line, names = c("Tr1 axe1", "Tr1 axe2", "Tr1 axe3", "Tr1 axe4"), main = "Correlation between trait 1 and axes for line species (res)", ylim = c(0,1))
boxplot(tr1x1_line, tr2x2_line, tr2x3_line, tr2x4_line, names = c("Tr2 axe1", "Tr2 axe2", "Tr2 axe3", "Tr2 axe4"), main = "Correlation between trait 2 and axes for line species (res)", ylim = c(0,1))

boxplot(tr1x1_col, tr1x2_col, tr1x3_col, tr1x4_col, names = c("Tr1 axe1", "Tr1 axe2", "Tr1 axe3", "Tr1 axe4"), main = "Correlation between trait 1 and axes for column species (con)", ylim = c(0,1))
boxplot(tr1x1_col, tr2x2_col, tr2x3_col, tr2x4_col, names = c("Tr2 axe1", "Tr2 axe2", "Tr2 axe3", "Tr2 axe4"), main = "Correlation between trait 2 and axes for column species (con)", ylim = c(0,1))

```

```{r}
s.corcircle(cor(simulated_network$trait$resource_giant_compo, afc.tab$li))
```


# Take care of the weird axes

```{r}
nb_frame = 20
spe_resource = 50
spe_consumer = 50
simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                   nb_location = nb_frame,
                                   know_env_grad_pos = T, ninter = 5000, 
                                   mean_tol = 0.1, sd_tol = 0.1,
                                   mean_tol_env = 0.1, sd_tol_env = 0.1,
                                   delta = 1, corr_env_trait = 0,
                                   ratio_grad = 0.7)

not_giant_compo <- connected_compo(simulated_network, nb_frame)
simulated_network <- filter_giant_component(simulated_network, not_giant_compo)


tab = Reduce('+', simulated_network$obs_giant_compo) # afin de générer le réseau moyen
afc.tab=dudi.coa(tab, nf = 18, scannf = FALSE)

(corr_line <- cor(simulated_network$trait$resource_giant_compo, afc.tab$li)) # resource correlation with the rows of the CA depending on the axis
(corr_col <- cor(simulated_network$trait$consumer_giant_compo[,1,], afc.tab$co)) # consumer correlation with the columns of the CA depending on the axis

plot(RV.randtest(data.frame(simulated_network$trait$resource_giant_compo), afc.tab$li))
plot(RV.randtest(data.frame(simulated_network$trait$consumer_giant_compo), afc.tab$co))

s.corcircle(corr_line)
s.corcircle(corr_col)
ade4::table.value(tab, clegend = 0)
scatter(afc.tab)


centrality <- closeness(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0]))

sort(centrality)
# color depending on the centrality
fine = 10 # this will adjust the resolving power.
pal = colorRampPalette(c('red','green'))

#this gives you the colors you want for every point
graphCol = pal(fine)[as.numeric(cut(log(centrality),breaks = fine))]

# now you just need to plot it with those colors
plot(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0]), vertex.size=5, vertex.color=graphCol)

# plot(graph_from_incidence_matrix(simulated_network$obs_giant_compo[[1]][]))

# for (i in 1:nb_frame){
#   print(simulated_network$obs_giant_compo[[i]]["res33", "con17"])
# }

# closeness(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0]))
# sort(closeness(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0])))

which(degree(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0]))>5)
degree(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0]))>5

fou1 = foucart(lapply(simulated_network$obs_giant_compo, as.data.frame), scannf = F, nf = 5)
plot(fou1, xax = 1)

# boxplot.stats(afc.tab$l1[,"RS2"])
ax_position <- afc.tab$l1[,"RS1"]
# which(ax_position < quantile(ax_position, 0.025))
# which(ax_position > quantile(ax_position, 0.975))
which(ax_position < mean(ax_position) - 3 * sd(ax_position))
which(ax_position >mean(ax_position) + 3 * sd(ax_position))
afc.tab$l1
hist(ax_position)
quantile(ax_position)

ax_position <- afc.tab$c1[,"CS1"]
# which(ax_position < quantile(ax_position, 0.025))
# which(ax_position > quantile(ax_position, 0.975))
which(ax_position < mean(ax_position) - 3 * sd(ax_position))
which(ax_position >mean(ax_position) + 3 * sd(ax_position))
afc.tab$c1
hist(ax_position)
quantile(ax_position)

cluster_spe <- cluster_edge_betweenness(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0]))
dendPlot(cluster_spe)

barplot(fou1$eig)
table.value(ifelse(tab> 0, 1,0), ppoint.cex=0.5)
```

```{r}

```


```{r}
s.arrow(diag(sqrt(fou1$cw)) %*% as.matrix(fou1$c1), label = rownames(fou1$l1))
s.label(fou1$li, add.plot = TRUE)

## à faire avec foucart

cor(diag(sqrt(fou1$cw)) %*% as.matrix(fou1$c1), simulated_network$trait$resource_giant_compo)

hist(simulated_network$trait$resource_giant_compo)

## second
s.arrow(diag(sqrt(fou1$lw)) %*% as.matrix(fou1$l1), label = rownames(fou1$l1))
s.label(fou1$co, add.plot = TRUE)
```


```{r}
transitivity(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0]), type = "barrat")
```

```{r}
cluster_spe <- cluster_louvain(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0]))
cluster_spe$membership
palette <- rainbow(max(cluster_spe$membership))
plot(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0]), vertex.size=5, vertex.color = palette[cluster_spe$membership])
```
```{r}
cluster_spe <- cluster_edge_betweenness(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0]))
dendPlot(cluster_spe)
cluster_spe$membership
palette <- rainbow(max(cluster_spe$membership))
plot(graph_from_incidence_matrix(tab[rowSums(tab) != 0, colSums(tab) != 0]), vertex.size=5, vertex.color = palette[cluster_spe$membership])
dendPlot(cluster_spe)
```

```{r}
spe
```


```{r}
print("for consummer")
vert_a = names(which(degree(graph_from_incidence_matrix(tab)) == 1))
vert_a_con <- grep("con", vert_a, value = T)
conn_vert_b_con <- c()
vert_b_res <- c()
vert_a_b_con <- c()
for(a in vert_a_con){
  vert_a_b_con <- c(vert_a_b_con, tab[which(tab[,a] !=0), a])
  conn_vert_b_con <- c(conn_vert_b_con, rowSums(tab)[names(which(tab[,a] !=0))])
}
vert_a_con
vert_a_b_con
conn_vert_b_con


print("for resource")
vert_a = names(which(degree(graph_from_incidence_matrix(tab)) == 1))
vert_a_res <- grep("res", vert_a, value = T)
conn_vert_b_res <- c()
vert_b_res <- c()
vert_a_b_res <- c()
for(a in vert_a_res){
  vert_a_b_res <- c(vert_a_b_res, tab[a, which(tab[a,] !=0)])
  conn_vert_b_res <- c(conn_vert_b_res, colSums(tab)[names(which(tab[a,] !=0))])
}
vert_a_res
vert_a_b_res
conn_vert_b_res

```

```{r}
filt_netw <- rep(list(NA), nb_frame)
for (i in 1:nb_frame) {
  filt_netw[[i]] <- simulated_network$obs_giant_compo[[1]][,-which(colnames(simulated_network$obs_giant_compo[[1]]) == "con17")]
}
# filt_netw
fou1 = foucart(lapply(filt_netw, as.data.frame), scannf = F, nf = 5)
plot(fou1)
```


# correlation trait matching reconstruit et th

```{r}
rv_distrib <- c()
for (i in 1:30) {
  nb_frame = 10
  spe_resource = 50
  spe_consumer = 50
  simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                     nb_location = nb_frame,
                                     know_env_grad_pos = T, ninter = 5000, 
                                     mean_tol = 0.1, sd_tol = 0.1,
                                     mean_tol_env = 0.1, sd_tol_env = 0.1,
                                     delta = 1, corr_env_trait = 0,
                                     ratio_grad = 0.7)
  
  not_giant_compo <- connected_compo(simulated_network, nb_frame)
  simulated_network <- filter_giant_component(simulated_network, not_giant_compo)
  
  
  tab = Reduce('+', simulated_network$obs_giant_compo) # afin de générer le réseau moyen
  afc.tab=dudi.coa(tab, nf = 18, scannf = FALSE)
  
  toto = reconst_afc(afc.tab, nf = 4)
  
  # plot(RV.randtest(toto, data.frame(simulated_network$trait$trait_mat_giant_compo), nrepet = 999))
  
  rv <- RV.randtest(toto, data.frame(simulated_network$trait$trait_mat_giant_compo), nrepet = 100)
  
  rv_distrib <- c(rv_distrib, rv$obs)
  # rv$obs 
}
hist(rv_distrib)
median(rv_distrib)
```


# For différent parameters, try out the correlation between the reconsytructed axes and the theoritical traits

```{r}
num_simu = 10
tr1ax1 <- c()
tr2ax2 <- c()
rv_distrib <- c()
ratio_grad = 0.7
mean_tol_env = 0.1
ninter = 2500
nb_frame = 2

for (i in 1:50) {
  spe_resource = 50
  spe_consumer = 50
  simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                     nb_location = nb_frame,
                                     know_env_grad_pos = T, ninter = ninter, 
                                     mean_tol = 0.1, sd_tol = 0.1,
                                     mean_tol_env = mean_tol_env, sd_tol_env = 0.1,
                                     delta = 1, corr_env_trait = 0,
                                     ratio_grad = ratio_grad)
  
  not_giant_compo <- connected_compo(simulated_network, nb_frame)
  simulated_network <- filter_giant_component(simulated_network, not_giant_compo)
  
  
  tab = Reduce('+', simulated_network$obs_giant_compo) # afin de générer le réseau moyen
  afc.tab=dudi.coa(tab, nf = 18, scannf = FALSE)
  
    
  corr_line <- cor(simulated_network$trait$resource_giant_compo, afc.tab$li) # resource correlation with the rows of the CA depending on the axis
  corr_col <- cor(simulated_network$trait$consumer_giant_compo[,1,], afc.tab$co) # consumer correlation with the columns of the CA depending on the axis
  tr1ax1 <- c(tr1ax1, corr_line[1,1])
  tr2ax2 <- c(tr2ax2, corr_line[2,2])
}
print(tr1ax1)
# jpeg(file= paste0("~/Images/trait_reconstruction_performance/simu",num_simu,"_1.jpeg"))
boxplot(abs(tr1ax1), ylim = c(0,1), main = paste0("Absolute correlation between trait 1 and ax 1 when the \n environement variance is ", mean_tol_env, " folds the gradient size (flat distrib) \n ratio_grad = ", ratio_grad, ", ", nb_frame, " frames, ", ninter, " obs each"))
# dev.off()

print(tr2ax2)
# jpeg(file= paste0("~/Images/trait_reconstruction_performance/simu",num_simu,"_2.jpeg"))
boxplot(abs(tr2ax2), ylim = c(0,1), main = paste0("Absolute correlation between trait 2 and ax 2 when the \n environement variance is ", mean_tol_env, " folds the gradient size (flat distrib) \n ratio_grad = ", ratio_grad, ", ", nb_frame, " frames, ", ninter, " obs each"))
# dev.off()

num_simu = num_simu + 1
```

```{r}
parameters <- c(0.01, 0.1, 1, 10, 100)
nb_rep = 30


param_tr1ax1 <- data.frame(matrix(ncol = length(parameters), nrow = nb_rep))
colnames(param_tr1ax1) <- as.character(parameters)

param_tr2ax2 <- data.frame(matrix(ncol = length(parameters), nrow = nb_rep))
colnames(param_tr2ax2) <- as.character(parameters)

for (param in parameters) {
  tr1ax1 <- c()
  tr2ax2 <- c()
  rv_distrib <- c()
  ratio_grad = 0.7
  # mean_tol_env = 0.1
  mean_tol_env = param
  ninter = 2500
  nb_frame = 2
  
  for (i in 1:nb_rep) {
    spe_resource = 50
    spe_consumer = 50
    simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                       nb_location = nb_frame,
                                       know_env_grad_pos = T, ninter = ninter, 
                                       mean_tol = 0.1, sd_tol = 0.1,
                                       mean_tol_env = mean_tol_env, sd_tol_env = 0.1,
                                       delta = 0.1, corr_env_trait = 0,
                                       ratio_grad = ratio_grad,
                                       magn_res_max = 20000)
    
    not_giant_compo <- connected_compo(simulated_network, nb_frame)
    simulated_network <- filter_giant_component(simulated_network, not_giant_compo)
    
    
    tab = Reduce('+', simulated_network$obs_giant_compo) # afin de générer le réseau moyen
    afc.tab=dudi.coa(tab, nf = 10, scannf = FALSE)
    
      
    corr_line <- cor(simulated_network$trait$resource_giant_compo, afc.tab$li) # resource correlation with the rows of the CA depending on the axis
    corr_col <- cor(simulated_network$trait$consumer_giant_compo[,1,], afc.tab$co) # consumer correlation with the columns of the CA depending on the axis
    tr1ax1 <- c(tr1ax1, corr_line[1,1])
    tr2ax2 <- c(tr2ax2, corr_line[2,2])
  }
  param_tr1ax1[,as.character(param)] <- abs(tr1ax1)
  param_tr2ax2[,as.character(param)] <- abs(tr2ax2)
  
}
# param_tr1ax1
boxplot(param_tr1ax1, ylim = c(0, 1), main = "Optimum trait 1 ax 1 correlation depending on mean tol env")
# param_tr2ax2
boxplot(param_tr2ax2, ylim = c(0, 1),main = "Optimum trait 2 ax 2 correlation depending on mean tol env")
```

The correlation decreases when the mean_tol_env gets too high because, the consumers are gathered across a bigger niche and hence, there is less and maybe a sampling effect, we just need to increase magn_res_max


```{r}
parameters <- c(250, 2500, 25000)
nb_rep = 30


param_tr1ax1 <- data.frame(matrix(ncol = length(parameters), nrow = nb_rep))
colnames(param_tr1ax1) <- as.character(parameters)

param_tr2ax2 <- data.frame(matrix(ncol = length(parameters), nrow = nb_rep))
colnames(param_tr2ax2) <- as.character(parameters)

for (param in parameters) {
  tr1ax1 <- c()
  tr2ax2 <- c()
  rv_distrib <- c()
  ratio_grad = 0.7
  mean_tol_env = 1
  # ninter = 2500
  ninter = param
  nb_frame = 2
  
  for (i in 1:nb_rep) {
    spe_resource = 50
    spe_consumer = 50
    simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                       nb_location = nb_frame,
                                       know_env_grad_pos = T, ninter = ninter, 
                                       mean_tol = 0.1, sd_tol = 0.1,
                                       mean_tol_env = mean_tol_env, sd_tol_env = 0.1,
                                       delta = 1, corr_env_trait = 0,
                                       ratio_grad = ratio_grad)
    
    not_giant_compo <- connected_compo(simulated_network, nb_frame)
    simulated_network <- filter_giant_component(simulated_network, not_giant_compo)
    
    
    tab = Reduce('+', simulated_network$obs_giant_compo) # afin de générer le réseau moyen
    afc.tab=dudi.coa(tab, nf = 10, scannf = FALSE)
    
      
    corr_line <- cor(simulated_network$trait$resource_giant_compo, afc.tab$li) # resource correlation with the rows of the CA depending on the axis
    corr_col <- cor(simulated_network$trait$consumer_giant_compo[,1,], afc.tab$co) # consumer correlation with the columns of the CA depending on the axis
    tr1ax1 <- c(tr1ax1, corr_line[1,1])
    tr2ax2 <- c(tr2ax2, corr_line[2,2])
  }
  param_tr1ax1[,as.character(param)] <- abs(tr1ax1)
  param_tr2ax2[,as.character(param)] <- abs(tr2ax2)
  
}
# param_tr1ax1
boxplot(param_tr1ax1, ylim = c(0, 1), main = "Optimum trait 1 ax 1 correlation depending on \n the   number of observations per frames")
# param_tr2ax2
boxplot(param_tr2ax2, ylim = c(0, 1),main = "Optimum trait 2 ax 2 correlation depending on \n the number of observations per frames")
```

```{r}
parameters <- c(2, 4, 10, 20)
nb_rep = 30


param_tr1ax1 <- data.frame(matrix(ncol = length(parameters), nrow = nb_rep))
colnames(param_tr1ax1) <- as.character(parameters)

param_tr2ax2 <- data.frame(matrix(ncol = length(parameters), nrow = nb_rep))
colnames(param_tr2ax2) <- as.character(parameters)

for (param in parameters) {
  tr1ax1 <- c()
  tr2ax2 <- c()
  rv_distrib <- c()
  ratio_grad = 0.7
  mean_tol_env = 1
  ninter = 2500 / param
  nb_frame = param
  
  for (i in 1:nb_rep) {
    spe_resource = 50
    spe_consumer = 50
    simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                       nb_location = nb_frame,
                                       know_env_grad_pos = T, ninter = ninter, 
                                       mean_tol = 0.1, sd_tol = 0.1,
                                       mean_tol_env = mean_tol_env, sd_tol_env = 0.1,
                                       delta = 1, corr_env_trait = 0,
                                       ratio_grad = ratio_grad)
    
    not_giant_compo <- connected_compo(simulated_network, nb_frame)
    simulated_network <- filter_giant_component(simulated_network, not_giant_compo)
    
    
    tab = Reduce('+', simulated_network$obs_giant_compo) # afin de générer le réseau moyen
    afc.tab=dudi.coa(tab, nf = 10, scannf = FALSE)
    
      
    corr_line <- cor(simulated_network$trait$resource_giant_compo, afc.tab$li) # resource correlation with the rows of the CA depending on the axis
    corr_col <- cor(simulated_network$trait$consumer_giant_compo[,1,], afc.tab$co) # consumer correlation with the columns of the CA depending on the axis
    tr1ax1 <- c(tr1ax1, corr_line[1,1])
    tr2ax2 <- c(tr2ax2, corr_line[2,2])
  }
  param_tr1ax1[,as.character(param)] <- abs(tr1ax1)
  param_tr2ax2[,as.character(param)] <- abs(tr2ax2)
  
}
# param_tr1ax1
boxplot(param_tr1ax1, ylim = c(0, 1), main = "Optimum trait 1 ax 1 correlation depending on \n the   number of frames")
# param_tr2ax2
boxplot(param_tr2ax2, ylim = c(0, 1),main = "Optimum trait 2 ax 2 correlation depending on \n the number of frames")
```

```{r}
parameters <- c(0, 0.01, 0.1, 1, 2)
nb_rep = 30


param_tr1ax1 <- data.frame(matrix(ncol = length(parameters), nrow = nb_rep))
colnames(param_tr1ax1) <- as.character(parameters)

param_tr2ax2 <- data.frame(matrix(ncol = length(parameters), nrow = nb_rep))
colnames(param_tr2ax2) <- as.character(parameters)

for (param in parameters) {
  tr1ax1 <- c()
  tr2ax2 <- c()
  rv_distrib <- c()
  ratio_grad = 0.7
  mean_tol_env = 1
  ninter = 2500
  nb_frame = 2
  
  for (i in 1:nb_rep) {
    spe_resource = 50
    spe_consumer = 50
    simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                       nb_location = nb_frame,
                                       know_env_grad_pos = T, ninter = ninter, 
                                       mean_tol = 0.1, sd_tol = 0.1,
                                       mean_tol_env = mean_tol_env, sd_tol_env = 0.1,
                                       delta = param, corr_env_trait = 0,
                                       ratio_grad = ratio_grad,
                                       magn_res_max = 20000)
    
    not_giant_compo <- connected_compo(simulated_network, nb_frame)
    simulated_network <- filter_giant_component(simulated_network, not_giant_compo)
    
    
    tab = Reduce('+', simulated_network$obs_giant_compo) # afin de générer le réseau moyen
    afc.tab=dudi.coa(tab, nf = 10, scannf = FALSE)
    
      
    corr_line <- cor(simulated_network$trait$resource_giant_compo, afc.tab$li) # resource correlation with the rows of the CA depending on the axis
    corr_col <- cor(simulated_network$trait$consumer_giant_compo[,1,], afc.tab$co) # consumer correlation with the columns of the CA depending on the axis
    tr1ax1 <- c(tr1ax1, corr_line[1,1])
    tr2ax2 <- c(tr2ax2, corr_line[2,2])
  }
  param_tr1ax1[,as.character(param)] <- abs(tr1ax1)
  param_tr2ax2[,as.character(param)] <- abs(tr2ax2)
  
}
# param_tr1ax1
boxplot(param_tr1ax1, ylim = c(0, 1), main = "Optimum trait 1 ax 1 correlation depending on delta")
# param_tr2ax2
boxplot(param_tr2ax2, ylim = c(0, 1), main = "Optimum trait 2 ax 2 correlation depending on delta")
```

# correlation

```{r}
ratio_grad = 0.7
mean_tol_env = 0.1
ninter = 2500
nb_frame = 10

col_corr_mean <- c()
col_corr_median <- c()

for (i in 1:100) {
  spe_resource = 50
  spe_consumer = 50
  simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                     nb_location = nb_frame,
                                     know_env_grad_pos = T, ninter = ninter, 
                                     mean_tol = 0.1, sd_tol = 0.1,
                                     mean_tol_env = mean_tol_env, sd_tol_env = 0.1,
                                     delta = 1, corr_env_trait = 0,
                                     ratio_grad = ratio_grad)
  
  not_giant_compo <- connected_compo(simulated_network, nb_frame)
  simulated_network <- filter_giant_component(simulated_network, not_giant_compo)
  
  
  tab = Reduce('+', simulated_network$obs_giant_compo) # afin de générer le réseau moyen
  afc.tab=dudi.coa(tab, nf = 18, scannf = FALSE)
  
    
  corr_line <- cor(simulated_network$trait$resource_giant_compo, afc.tab$li) # resource correlation with the rows of the CA depending on the axis
  corr_col <- cor(simulated_network$trait$consumer_giant_compo[,1,], afc.tab$co) # consumer correlation with the columns of the CA depending on the axis
  tr1ax1 <- c(tr1ax1, corr_line[1,1])
  tr2ax2 <- c(tr2ax2, corr_line[2,2])
  
  reconst_trait_match = reconst_afc(afc.tab, nf = 12) # pour reconstruire le réseau moyen à partir des positions sur l'AFC mais on voudrait plustôt obtenir le trait matching
  
  # ade4::table.value(reconst_trait_match, x = rank(afc.tab$co[,1]), y = rank(afc.tab$li[,1]), csize = 0.8)
  
  column_corr <- sapply(1:ncol(reconst_trait_match), function(i) {
    cor(reconst_trait_match[,i], simulated_network$trait$trait_mat_giant_compo[,i])
  })
  col_corr_median <- c(col_corr_median, median(column_corr))
  col_corr_mean <- c(col_corr_mean, mean(column_corr))
  # boxplot(abs(column_corr), ylim = c(0,1))
  # print(which(is.na(column_corr)))
}

boxplot(col_corr_mean, ylim = c(0,1))
boxplot(col_corr_median, ylim = c(0,1))
```


# Tester empiriquement reconstruction traits depuis trait matching th

```{r}
ratio_grad = 0.7
mean_tol_env = 0.1
ninter = 2500
nb_frame = 10

spe_resource = 50
spe_consumer = 50
simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                   nb_location = nb_frame,
                                   know_env_grad_pos = T, ninter = ninter, 
                                   mean_tol = 0.7, sd_tol = 0.1,
                                   mean_tol_env = mean_tol_env, sd_tol_env = 0.1,
                                   delta = 1, corr_env_trait = 0,
                                   ratio_grad = ratio_grad)

not_giant_compo <- connected_compo(simulated_network, nb_frame)
simulated_network <- filter_giant_component(simulated_network, not_giant_compo)

ade4::table.value(simulated_network$trait$matching_matrix)

############################# avec log ###########################################
# 
# afc.tab=dudi.coa(log(simulated_network$trait$matching_matrix + 1), nf = 4, scannf = F)
# # afc.tab=dudi.coa(simulated_network$trait$matching_matrix, nf = 4, scannf = F)
# 
# reconst_trait_match = reconst_afc(afc.tab, nf = 4) 
# 
# ade4::table.value(reconst_trait_match)
# 
# column_corr <- sapply(1:ncol(reconst_trait_match), function(i) {
#     cor(reconst_trait_match[,i], simulated_network$trait$matching_matrix[,i])
#   })
# 
# boxplot(column_corr, ylim = c(0, 1))
# 
# corr_line <- cor(simulated_network$trait$trait_resource, afc.tab$li) # resource correlation with the rows of the CA depending on the axis
# corr_col <- cor(simulated_network$trait$trait_consumer[,1,], afc.tab$co) # consumer correlation with the columns of the CA depending on the axis
# 
# print(corr_col)
# print(corr_line)
# 
# scatter(afc.tab)
# 
# # log(matrix(c(1:25), ncol = 5, nrow = 5)) + 1

################################ sans log ###########################################

afc.tab=dudi.coa(simulated_network$trait$matching_matrix, nf = 4, scannf = F)

reconst_trait_match = reconst_afc(afc.tab, nf = 4) 

ade4::table.value(reconst_trait_match)

column_corr <- sapply(1:ncol(reconst_trait_match), function(i) {
    cor(reconst_trait_match[,i], simulated_network$trait$matching_matrix[,i])
  })

boxplot(column_corr, ylim = c(0, 1))

corr_line <- cor(simulated_network$trait$trait_resource, afc.tab$li) # resource correlation with the rows of the CA depending on the axis
corr_col <- cor(simulated_network$trait$trait_consumer[,1,], afc.tab$co) # consumer correlation with the columns of the CA depending on the axis

print(corr_col)
print(corr_line)

scatter(afc.tab)

# log(matrix(c(1:25), ncol = 5, nrow = 5)) + 1
```



## effet des parametres sur l

```{r}

```




# representation with bipartite

```{r}
library(bipartite)

ratio_grad = 0.7
mean_tol_env = 1
ninter = 1000
nb_frame = 10

col_corr_mean <- c()
col_corr_median <- c()

spe_resource = 100
spe_consumer = 100
simulated_network <- env_grad_netw(nb_resource = spe_resource , nb_consumer = spe_consumer,
                                   nb_location = nb_frame,
                                   know_env_grad_pos = T, ninter = ninter, 
                                   mean_tol = 0.1, sd_tol = 0.1,
                                   mean_tol_env = mean_tol_env, sd_tol_env = 0.1,
                                   delta = 0, corr_env_trait = 0,
                                   ratio_grad = ratio_grad)

not_giant_compo <- connected_compo(simulated_network, nb_frame)
simulated_network <- filter_giant_component(simulated_network, not_giant_compo)

tab = Reduce('+', simulated_network$obs_giant_compo) # afin de générer le réseau moyen

# visweb(simulated_network$obs_giant_compo[[1]])

for(i in 1:nb_frame){
  plotweb(simulated_network$obs_giant_compo[[i]])
}

visweb(tab, type = "diagonal")
# plotweb(tab)
```

```{r}
visweb(matrix(runif(900, 0, 10), nrow = 30, ncol = 30), type = "diagonal")
```


# calcul hypervolume

```{r}
library(cxhull)
points <- rbind(
c(0.5,0.5,0.5),
c(0,0,0),
c(0,0,1),
c(0,1,0),
c(0,1,1),
c(1,0,0),
c(1,0,1),
c(1,1,0),
c(1,1,1)
)
hull <- cxhull(points)

hull$volume
```

