---
title: "data_vincent"
author: "Rémi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(igraph)
```


The R dataset consists of the bipartite plant-pollinator interaction matrices (*bipartite.networks*) for 18 subseas (3 subseasons per 6 years), 18 subseasonal dependence matrices (*directed.networks*), and the plant species (*plant.id*) and pollinator species (*pollinator.id*) identification codes. The code used for the analyses (miele-JAE.R) is also provided.

18 seasons, 59 plants and 196 pollinators

```{r}
season_1 <- bipartite.networks[1, , ]
# View(season_1)
# which(season_1[2,] != 0)# pas beaucoup differents de 0

# to keep only the connected species
test <- season_1[which(rowSums(season_1) != 0), which(colSums(season_1) != 0)]
dim(test)
# test

# plot(graph_from_incidence_matrix(test), vertex.size=5)

# str(bipartite.networks)
```

```{r}
for (i in 1:18) {
  sampled_season <- bipartite.networks[i, , ]
  active_sample <- sampled_season[which(rowSums(sampled_season) != 0), which(colSums(sampled_season) != 0)]
  print(dim(active_sample))
}
```


```{r}
pollinator.id
plant.id
```

